---
title: "TidyTuesday-2022-Wk-17-OnlineSportingGoods"
author: "Philip Wong"
date: '`r Sys.Date()`'
output: rmarkdown::github_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../Output") })
always_allow_html: yes
---

<strong> Week 17 - TidyTuesday Challenge: </strong> <br> 
This week we will be analyzing a ***product-related*** dataset from Kaggle as opposed to the datasets provided in TidyTuesday's GitHub. <br>   
***Project Source:*** https://www.kaggle.com/datasets/podsyp/how-to-do-product-analytics <br>

***Project Scenario*** <br>
Online store of sporting goods: clothing, shoes, accessories and sports nutrition. <br>

On the main page of the store's website, users will view banners in order to stimulate their sales. Each user will be able to see one of 5 banners, randomly selected and displayed. Each banner either advertises a specific product or the entire company. We believe the experience with banners can vary by segment, and their effectiveness may depend on the characteristics of user behavior." <br>

***Problem:*** <br>
Let's use our analytical skills to drive decisions around sales/user growth! We're looking to answer the following questions. <br>
      - Which banner contains the highest number of returning visitors?  <br>
      - Which days of the week results in the highest average user traffic (clicks & impressions)? <br> 
      - What time of day results in the highest average user traffic, given same timezone? <br> 
      - Which banner drove the most sales? Which banner should we remove? <br>
      - For each banner, what's the average timespan before users make their first purchase? <br> 
      
### [1] Import Packages & Datasets

```{r, echo=TRUE, message=FALSE}
library(tidyverse)
library(data.table)
library(scales)
library(reshape2)
library(ggcorrplot) # to build correlation heatmap
library(knitr)
library(lubridate)
library(plotly)
library(zoo)

knitr::opts_chunk$set(cache = TRUE, 
                      warning = FALSE,
                      message = FALSE, 
                      echo = TRUE, 
                      fig.width = 13, 
                      fig.height = 5)

product <- read_csv("../Data/TTWK17_product.csv")

theme_blue <- "#2E15DA"
theme_grey <- "#858286" 
theme_purple <- "#A714E7"
```

### [2] Data exploration & pre-processing

##### 2.1 - Dataframe summary

***Data Description:*** <br>

```{r}
# let's view the dataframe size & datatypes of each variable
glimpse(product)
```

##### 2.2 - Verify any NULL values

```{r}
print(colSums(is.na(product)))
```
##### 2.3 - Data manipulation

```{r}
product_df <- product %>% 
  mutate(hour_of_day = as.numeric(format(time, format = "%H"))) %>% 
  mutate_if(is.POSIXt, as.Date) %>% 
  rename(date = "time",
         banner = "product",
         device = "site_version",
         event = "title") %>% 
  mutate(day_of_week = weekdays(date)) %>%
  mutate(week_start_date = as.Date(cut(as.Date(date), "week"), "%Y-%m-%d")-1) %>% 
  mutate(month_start_date = as.Date(as.yearmon(date, format = "%b%Y"), frac = 0)) %>% 
  group_by(device, banner, user_id) %>% 
  mutate(first_session = case_when(date == min(date) ~ 1, TRUE ~ 0)) %>% 
  ungroup() %>% 
  mutate(event = case_when(event == "banner_show" ~ 'impressions',
                           event == "banner_click" ~ 'clicks',
                           event == "order" ~ 'orders',
                           TRUE ~ event))

head(product_df)
```

Within our dataset, we have information about which banners appeared to the users, whether they clicked on it, as well as information about their user purchases. <br>
  - order_id - unique purchase number (NA for banner clicks and impressions) <br>
  - user_id - unique identifier of the client <br>
  - page_id - unique page number for event bundle (NA for purchases) <br>
  - banner - 5 unique products/groups <br>
  - device - mobile or desktop <br>
  - time - time of the action <br>
  - title - type of event (show, click or purchase) <br>
  - target - target class <br>

### [3] Perform data analysis

##### 3.1 - Which banner contains the highest number of returning visitors?

```{r}
plot1 <- product_df %>% 
  group_by(banner, user_id) %>% 
  mutate(is_returning_user = case_when(n_distinct(date) > 1 ~ 1, 
                                     TRUE ~ 0)) %>% 
  ungroup() %>% 
  group_by(banner) %>% 
  summarise(returning_user_count = n_distinct(user_id[is_returning_user == 1])) %>% 
  ungroup() 

plot1 %>% 
  ggplot(aes(x = reorder(banner, returning_user_count), y = returning_user_count)) +
  geom_bar(stat = 'identity', fill = theme_blue, alpha = 0.9) +
  theme_bw() +
  xlab("banner") +
  ylab("number of returning users")
```

##### 3.2 - Which days of the week results in the highest average user traffic (clicks & impressions)?

```{r}
plot2 <- product_df %>% 
  filter(event %in% c("clicks","impressions")) %>%
  group_by(day_of_week, event) %>% 
  summarise(avg_user_traffic = round(n() / n_distinct(week_start_date), 0)) %>% 
  ungroup()

ordered_weekDays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
  
plot2 %>% 
  ggplot(aes(x = factor(day_of_week, level = ordered_weekDays), y = avg_user_traffic, fill = event)) +
  geom_bar(stat = 'identity') +
  theme_bw() +
  xlab("day of week") +
  ylab("average user traffic") +
  scale_fill_manual(values = c("clicks" = theme_grey, "impressions" = alpha(theme_blue, 0.9)))
```

##### 3.3 - What time of day results in the highest average user traffic, given same timezone?

```{r}
plot3 <- product_df %>% 
  group_by(hour_of_day, event) %>%
  summarise(avg_user_traffic = round(n() / n_distinct(date), 0)) %>% 
  ungroup()
  
plot3 %>% 
  ggplot(aes(x = as.factor(hour_of_day), y = avg_user_traffic, fill = event)) +
  geom_bar(stat = 'identity') +
  theme_bw() +
  xlab("hour of day") +
  ylab("average user traffic") +
  facet_wrap(~event, scales = "free") +
  theme(legend.position="none") +
  scale_fill_manual(values=c(alpha(theme_blue, 0.9),
                             theme_grey,
                             alpha(theme_purple, 0.9)))
```

##### 3.4 - Which banner drove the highest quantity of sales?

```{r}
plot4 <- product_df %>% 
  filter(event == 'orders') %>% 
  group_by(banner) %>% 
  summarise(total_orders = n()) %>% 
  ungroup()

plot4 %>% 
  ggplot(aes(x = reorder(banner, total_orders), y = total_orders)) +
  geom_bar(stat = 'identity', fill = alpha(theme_blue, 0.9)) +
  coord_flip() +
  theme_bw() +
  xlab("banner") +
  ylab("total number of orders")
```

##### 3.5 - For each banner, what's the average timespan before users make their first purchase?

```{r}
# first session by device, banner, and user
min_date <- product_df %>%
  filter(event != "orders") %>%
  group_by(device, banner, user_id) %>%
  summarise(first_session_date = min(date)) %>%
  ungroup()

# first order date by device, banner, and user
order_date <- product_df %>%
  filter(event == "orders") %>%
  group_by(device, banner, user_id) %>%
  summarise(order_date = min(date)) %>%
  ungroup()

min_date %>%
  left_join(order_date, by = c("device", "banner", "user_id")) %>%
  drop_na(order_date) %>%
  filter(order_date > first_session_date) %>%
  mutate(timespan = as.numeric(difftime(order_date, first_session_date, units = "days"))) %>% 
    ggplot(aes(x = banner, y = timespan, fill = device)) +
    geom_boxplot() +
    theme_bw() +
    facet_wrap(~device) +
    labs(x = "banner", y = "number of days to make a purchase") +
    theme(axis.text.x = element_text(angle = 90),
          legend.position="none") +
    scale_fill_manual(values=c(alpha(theme_blue, 0.9),
                                alpha(theme_grey, 0.9)))
```

